Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-05 Manuscript/manuscript.tex

\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
The dehydrated complexes of PS headgroup and calcium ions can also lead to the
phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
%
CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
the same force field without the NBfix parameters together with the NMR chemical shifts and
REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.

@@

%%% Sum up what is to be presented briefly
%The importance of treating polarizability for interactions of phospholipids with monovalent ions was demonstrated in the previous section. 
Electronic polarization is a non-negligible contribution to the interactions of calcium even in simple aqueous solutions of \ce{CaCl2} \citep{martinek17, kohagen16, Pluharova2014}. 
In this section,
we will show the results from simulations of neutral and negatively charged phospholipid bilayers at varying \ce{CaCl2} concentrations
using the recently developed models ECC-lipids and ECC-ions \citep{melcr18, martinek17}, 
which implicitly include the effects of electronic polarization through electronic continuum correction \citep{leontyev11}. 
Validated with the concept of a lipid electrometer in section \ref{section:electrometer_exp_sim},
these implicitly polarizable models yield accurate description of 
the interaction of \ce{Ca^{2+}} with both neutral and negatively charged phospholipids.

The changes of the head group order parameters $S^\alpha$ and $S^\beta$ from simulations and experiments 
are shown in Fig.~\ref{fig:delta_ordPar_CaCl} for a neutral POPC bilayer, 
and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS} for a negatively charged bilayer with the composition 5\,PC:1\,PS. 
For a direct comparison and a connection to our works \citep{catte16, nmrlipids_proj4},
we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
This also highlights the improvements in ECC-lipids over Lipid17 arising from the electronic polarization. 
%Although Lipid17 already belongs to the top-performing models in terms of the responses of the head group order parameters in those studies,  
%including electronic polarizability to form ECC-lipids improves the results even further.
The effect is probably the most striking for POPS, 
for which also the structure of a pure POPS bilayer with only counterions is dramatically improved with the augmentation.

%%% Discuss the changes of OPs and vectors from neutral and neg. membranes
%While the changes induced by \ce{CaCl2} are \emph{qualitatively} correct for Lipid14/17, 
Increasing concentrations of \ce{CaCl2} induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ in POPC. 
Although the total magnitude of the response of the PC head group order parameters 
is only slightly higher in the negatively charged bilayers than in the neutral bilayers, 
the shape of the changes in the latter shows a steeper onset at low concentrations. 
This is apparently due to the presence of POPS, 
which has a higher affinity to \ce{Ca^{2+}} compared to POPC. 
ECC-lipids is the first model,
which achieves a \emph{quantitative} agreement with the changes induced by \ce{CaCl2} in experiments.

\todo{SAMULI: The new results show that we cannot actually reproduce the rapid changes with low CaCl2
  concentrations in POPS order parameters. This should be somehow discussed.}

@@

The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
are summarized in Table~\ref{tab:binding}.
Even though the negatively charged membrane contains only 18\% of POPS, 
approximately 41\% of the total population of bound calcium cations is in contact with PS lipids
with 8\% bound only to them. 
This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
POPC interacts with the calcium cations almost entirely through its phosphate group 
in both neutral and negatively charged membranes. 
Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
however, they are always accompanied by interactions with phosphate groups. 
%but a very small negligable amount much below 1\%. 
\todo{Also COO- vs. phosphate binding should be discussed and the results should be
  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }

>>>>

@@

The interaction of various cations including calcium 
with  neutral and also negatively charged lipid bilayers
was thoroughly compared between simulations and experiments 
from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
in the references~\citenum{catte16, nmrlipids_proj4}. 
It was concluded that the interaction of calcium with zwitterionic PC bilayers 
is generally overestimated in all studied models \cite{catte16},
while none of the models was capable of interpreting the NMR structural parameters
with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
Accounting for the electronic polarization was identified as 
a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}. 
Here, we demonstrate that the same approach 
significantly improves the description of calcium binding 
also to the negatively charged PC:PS (5:1) bilayers. 
In correspondence with the Ref.~\citenum{nmrlipids_proj4},
we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
in PC and PS lipids 
to quantify the binding affinity of the calcium cations
and the induced structural changes
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
and to mark the improvement after treating electronic polarizability, 
we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
While the response of the order parameters of PC to \ce{CaCl2} concentrations
is qualitatively correct in the Lipid14/17 model \cite{lipid17},
it is quantitatively comparable to the experimental trends in ECC-lipids. 
Moreover, the structure of PS head group is fundamentaly improved
after accounting for the electronic polarization 
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).

Increasing concentrations of \ce{CaCl2} 
induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ 
in both neutral POPC bilayers \cite{catte16, melcr18}
as well as in the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4, roux90}. 
The change of the head group order parameters 
are well reproduced only by ECC-lipids in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.

The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
are very challenging to interpret from both experiments \cite{roux90} 
and also from MD simulations \cite{nmrlipids_proj4}. 
The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
are not found to follow the experimental profiles, 
nor any common trend,
with the currently available models \cite{nmrlipids_proj4}. 
In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
we plot these structural parameters for one of the models in that work, Lipid17, 
and for ECC-lipids. 
Although the agreement between ECC-lipids and experiments is far from optimal,
the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
are well captured by the model. 
For example the order parameter $\beta$ in POPS 
adopts a saturated value in experiments and ECC-lipids,
but it is growing with the \ce{CaCl2} concentration in Lipid17.
In addition, 
the forking of the order parameters $\alpha$ 
is almost constant in the experiment and ECC-lipids,
but it varies greatly in th Lipid17.

The changes of the order parameters $\alpha$ and $\beta$ at low \ce{CaCl2} concentrations
have a steep onset at low concentrations in the negatively charged bilayers compared to neutral POPC bilayers. 
However, such a detail is not captured by ECC-lipids
indicating the necessity for further improvements.
The source of this discrepancy may arise 
from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
and also possibly from the approximate mean-field treatment of the electronic polarization.

@@

The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
are summarized in Table~\ref{tab:binding}.
Even though the negatively charged membrane contains only 18\% of POPS, 
approximately 41\% of the total population of bound calcium cations is in contact with PS lipids
with 8\% bound only to them. 
This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
POPC interacts with the calcium cations almost entirely through its phosphate group 
in both neutral and negatively charged membranes. 
Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
however, they are always accompanied by interactions with phosphate groups. 
%but a very small negligable amount much below 1\%. 
\todo{Also COO- vs. phosphate binding should be discussed and the results should be
  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
%
\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
The dehydrated complexes of PS headgroup and calcium ions can also lead to the
phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
%
CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
the same force field without the NBfix parameters together with the NMR chemical shifts and
REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
